Fast solvers for the high-order FEM simplicial de Rham complex \thanks{Submitted to the editors July XX, 2024. \funding{ PDB and PEF were supported by EPSRC grant EP/W026260/1. } }

### 2024-06-17 paper/paper.tex
...

%------------------------------------------------------------------------------
% Header
%------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath, amssymb, gensymb, physics, upgreek, siunitx, bbm}
\usepackage{tikz, pgfplots, graphicx, epstopdf, float, hyperref, xspace}
\usepackage[caption=false]{subfig}
\usepackage{enumitem, verbatim, listings}
\usepackage{mathtools, stmaryrd}
\usepackage{booktabs, array, ragged2e}
\usepackage[export]{adjustbox}
\usepackage{cleveref}

@@

@@

\begin{abstract}
We present new high-order finite elements discretizing the $L^2$ de Rham
complex on triangular and tetrahedral meshes. The finite elements discretize
the same spaces as usual, but with different basis functions. They allow for
fast linear solvers based on static condensation and space decomposition
methods. The new elements build upon the definition of degrees of freedom
given in [L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz,
Comput.~Math.~Appl.,
39(7-8) (2000), pp.~29--38], and consist of integral moments
on an equilateral reference simplex with respect to a numerically computed
polynomial basis that is orthogonal in both the $L^2$- and
$H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
\mathrm{curl}, \mathrm{div}\}$). 
On the reference equilateral simplex, the
resulting stiffness matrix has diagonal interior block, and does not couple
together the interior and interface degrees of freedom. Thus, on the
reference simplex, the Schur complement resulting from elimination of
interior degrees of freedom is simply the interface block itself. This
sparsity is not preserved on arbitrary cells mapped from the reference cell.
Nevertheless, the interior-interface coupling is weak because it is only
induced by the geometric transformation. We devise a preconditioning
strategy by neglecting the interior-interface coupling. We precondition the
interface Schur complement with the interface block, and simply apply
point-Jacobi to precondition the interior block. The combination of this
approach with a space decomposition method on small subdomains constructed
around vertices and edges allows us to efficiently solve the
canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
$H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
counts that are robust with respect to the polynomial degree. 
\end{abstract}

@@

We discretize $\Ltwo$-de Rham complex on triangular and tetrahedral meshes.

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1}
\end{tikzcd}

\begin{tikzcd}
  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$-de Rham complex, and the finite element subcomplexes of the first and second kind.}
\end{figure}

>>>>

%------------------------------------------------------------------------------
% Header
%------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath, amssymb, gensymb, physics, upgreek, siunitx, bbm}
\usepackage{tikz, pgfplots, graphicx, epstopdf, float, hyperref, xspace}
\usepackage[caption=false]{subfig}
\usepackage{enumitem, verbatim, listings}
\usepackage{mathtools, stmaryrd}
\usepackage{booktabs, array, ragged2e}
\usepackage[export]{adjustbox}
\usepackage{cleveref}
\usepackage{todonotes}

@@

\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: \#1}}

@@

\begin{abstract}
We present new high-order finite elements discretizing the $L^2$ de Rham
complex on triangular and tetrahedral meshes. The finite elements discretize
the same spaces as usual, but with different basis functions. They allow for
fast linear solvers based on static condensation and space decomposition
methods. The new elements build upon the definition of degrees of freedom
given in [L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz,
Comput.~Math.~Appl.,
39(7-8) (2000), pp.~29--38], and consist of integral moments
on an equilateral reference simplex with respect to a numerically computed
polynomial basis that is orthogonal in both the $L^2$- and
$H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
\mathrm{curl}, \mathrm{div}\}$). 
On the reference equilateral simplex, the
resulting stiffness matrix has diagonal interior block, and does not couple
together the interior and interface degrees of freedom. Thus, on the
reference simplex, the Schur complement resulting from elimination of
interior degrees of freedom is simply the interface block itself. This
sparsity is not preserved on arbitrary cells mapped from the reference cell.
Nevertheless, the interior-interface coupling is weak because it is only
induced by the geometric transformation. We devise a preconditioning
strategy by neglecting the interior-interface coupling. We precondition the
interface Schur complement with the interface block, and simply apply
point-Jacobi to precondition the interior block. The combination of this
approach with a space decomposition method on small subdomains constructed
around vertices and edges allows us to efficiently solve the
canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
$H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
counts that are robust with respect to the polynomial degree, with a computational complexity of $\mathcal{O}(p^{6})$ flops in three dimensions
instead of the na\"ive $\mathcal{O}(p^{9})$.
\end{abstract}

@@

We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds.

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$-de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
\end{figure}
